### 2024-09-10 Article/main.tex
Some adjustments

\documentclass[a4paper, twoside,openright]{report}
\usepackage[T1]{fontenc} % Font encoding, T1 = it
\usepackage[utf8]{inputenc} % Input encoding - per caratteri particolari
\usepackage[english]{babel} % Lingua principale inglese
\usepackage{graphicx} % Per includere immagini esterne
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm]{geometry} %impaginazione e margini documento
\usepackage[fontsize=13pt]{scrextend} %dimensione font

@@

\begin{definition}[Comma category]\label{def:comma_category}
    Let $\cat{C \text{, } D \text{ and } E}$ be categories, and let $S: \cat{C \rightarrow E}$, $T:\cat{D \rightarrow E}$ be functors (source and target):
    \[
        \begin{tikzcd}
            \cat{C} \arrow[r, "S"] & \cat{E} & \cat{D} \arrow[l, "T" swap]
        \end{tikzcd}
    \]
    Then, the \emph{slice category $(S \downarrow T)$} is the category in which: 
    \begin{itemize}
        \item the objects are triples $(A, f, B)$, where $A \in \Ob(\cat{C})$, $B \in \Ob(\cat D)$ and $f: S(A) \rightarrow T(B)$ is an arrow of $\cat E$;
        \item the arrows are pairs $(c, d): (A, f, B) \rightarrow (C, g, D)$, where $c \in \Hom(\cat C)$ and $d \in \Hom(\cat D)$, such that the square below commutes;
        \[
            \begin{tikzcd}
            S(A) \arrow[r, "f"] \arrow[d, "S(c)" swap] & G(B) \arrow[d, "T(d)"] \\
            T(C) \arrow[r, "g" swap] & T(B)
            \end{tikzcd}
        \]
        \item composition of morphisms is obtained via pairwise composition, i.e., $(a, b) \circ (c, d) = (a \circ c, b \circ d)$.
    \end{itemize}
\end{definition}

@@

\begin{theorem}[Limit theorem]
    Let $\cat C$ be a category. Then $\cat C$ has all finite limits if and only if $\cat C$ has all finite products and all finite equalizers.
\end{theorem}

@@

@@

\begin{definition}\label{def:adh}[$\mathcal{M}$-adhesivity]
    Let $\cat C$ be a category and $\mathcal M$ a subclass of $\Mono(\cat C)$ containing all isomorphisms, closed under composition and stable under pullbacks and pushouts (Definition~\ref{def:stab_under_pb_po}).
    Then $\cat C$ is \emph{$\mathcal M$-adhesive} if
    \begin{enumerate}
        \item every cospan $C \xrightarrow[]{g} D \xleftarrow[]{m} B$ with $m \in \mathcal M$ can be completed to a pullback ( such pullbacks are called $\mathcal M$-pullbacks);
        \item every span $C \xleftarrow{m} A \xrightarrow{f} B$ with $ m \in \mathcal M$ can be completed to a pushout (such pushouts are called $\mathcal M$-pushouts);
        \item pushouts along $\mathcal M$-arrows are $\mathcal M$-Van Kampen squares.
    \end{enumerate}
    We also say that $\cat C$ is \emph{adhesive} when it is $\Mono(\cat C)$-adhesive, and \emph{quasiadhesive} when it is $\Reg(\cat C)$-adhesive.
\end{definition}

@@

@@

\begin{definition}[Quotient functor]
    The \emph{quotient functor} $Q: \mathbf{EqGrph \rightarrow Graph}$ is the functor that maps each graph with equivalence $\mathcal{G} = (G, \sim_G)$ to its \emph{quotient graph} $G/_{\sim_G} = (V_G/_{\sim_G}, E_G, s_G', t_G')$, with $s_G'([e]_{\sim_G}) = [v]_{\sim_G}$ if $s_G(e) = v$ and the same for $t_G'$, and each homomorphism between graphs with equivalences $h: \mathcal{G \rightarrow H}$ to $Q(h)$ defined by $Q(h_V)([v]_{\sim_G}) = [h_V(v)]_{\sim_H}$.
\end{definition}

>>>>

\documentclass[a4paper, twoside,openright]{report}
\usepackage[T1]{fontenc} % Font encoding, T1 = it
\usepackage[utf8]{inputenc} % Input encoding - per caratteri particolari
\usepackage[english]{babel} % Lingua principale inglese
\usepackage{graphicx} % Per includere immagini esterne
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm]{geometry} %impaginazione e margini documento
\usepackage[fontsize=12pt]{scrextend} %dimensione font

@@

\begin{definition}[Comma category]\label{def:comma_category}
    Let $\cat{C \text{, } D \text{ and } E}$ be categories, and let $S: \cat{C \rightarrow E}$, $T:\cat{D \rightarrow E}$ be functors (source and target):
    \[
        \begin{tikzcd}
            \cat{C} \arrow[r, "S"] & \cat{E} & \cat{D} \arrow[l, "T" swap]
        \end{tikzcd}
    \]
    Then, the \emph{comma category $(S \downarrow T)$} is the category in which: 
    \begin{itemize}
        \item the objects are triples $(A, f, B)$, where $A \in \Ob(\cat{C})$, $B \in \Ob(\cat D)$ and $f: S(A) \rightarrow T(B)$ is an arrow of $\cat E$;
        \item the arrows are pairs $(c, d): (A, f, B) \rightarrow (C, g, D)$, where $c \in \Hom(\cat C)$ and $d \in \Hom(\cat D)$, such that the square below commutes;
        \[
            \begin{tikzcd}
            S(A) \arrow[r, "f"] \arrow[d, "S(c)" swap] & G(B) \arrow[d, "T(d)"] \\
            T(C) \arrow[r, "g" swap] & T(B)
            \end{tikzcd}
        \]
        \item composition of morphisms is obtained via pairwise composition, i.e., $(a, b) \circ (c, d) = (a \circ c, b \circ d)$.
    \end{itemize}
\end{definition}

@@

\begin{theorem}\label{th:limit}[Limit theorem]
    Let $\cat C$ be a category. Then $\cat C$ has all finite limits if and only if $\cat C$ has all finite products and all finite equalizers.
\end{theorem}

@@

\begin{obs}\label{obs:limits_in_presh}
    So far we introduced categories of presheaves. In these categories, an interesting fact is that limits and colimits are computed pointwise -- i.e., the limit of a diagram in a category of presheaves is exactly the limit on each of its components.
\end{obs}

@@

\begin{definition}\label{def:adh}[$\mathcal{M}$-adhesivity]
    Let $\cat C$ be a category and $\mathcal M \subseteq \Mono(\cat C)$ containing all isomorphisms, closed under composition and stable under pullbacks and pushouts (Definition~\ref{def:stab_under_pb_po}).
    Then $\cat C$ is \emph{$\mathcal M$-adhesive} if
    \begin{enumerate}
        \item every cospan $C \xrightarrow[]{g} D \xleftarrow[]{m} B$ with $m \in \mathcal M$ can be completed to a pullback ( such pullbacks are called $\mathcal M$-pullbacks);
        \item every span $C \xleftarrow{m} A \xrightarrow{f} B$ with $ m \in \mathcal M$ can be completed to a pushout (such pushouts are called $\mathcal M$-pushouts);
        \item pushouts along $\mathcal M$-arrows are $\mathcal M$-Van Kampen squares.
    \end{enumerate}
    We also say that $\cat C$ is \emph{adhesive} when it is $\Mono(\cat C)$-adhesive, and \emph{quasiadhesive} when it is $\Reg(\cat C)$-adhesive.
\end{definition}

@@

These examples show how some kind of limits are computed among graphs, but we have already said something similar about categories of presheaves (Observation~\ref{obs:limits_in_presh}). In fact, The category of graphs in nothing than

@@

% \begin{definition}[Quotient functor]
%     The \emph{quotient functor} $Q: \mathbf{EqGrph \rightarrow Graph}$ is the functor that maps
%     each graph with equivalence $\mathcal{G} = (G, \sim_G)$ to its \emph{quotient graph} 
%     $G/_{\sim_G} = (V_G/_{\sim_G}, E_G, s_G', t_G')$, with $s_G'([e]_{\sim_G}) = [v]_{\sim_G}$ if 
%     $s_G(e) = v$ and the same for $t_G'$, and each homomorphism between graphs with equivalences 
%     $h: \mathcal{G \rightarrow H}$ to $Q(h)$ defined by $Q(h_V)([v]_{\sim_G}) = [h_V(v)]_{\sim_H}$.
% \end{definition}

% \begin{definition}[Quotient functor]
%     The \emph{quotient functor} $Q: \mathbf{EqGrph \rightarrow Graph}$ is the functor that maps each $\mathcal{G} = (G, \sim)$ onto its quotient graph $G/_{\sim} = (V_G/_{\sim},E_G, s_G', t_G')$, such that, for each edge $e$:
%     \begin{itemize}
%         \item $s_G'([e]_{\sim}) = [v]_{\sim_G}$
%     \end{itemize}
% \end{definition}

Some examples of limits will follow.

\begin{example}
    The terminal object in $\mathbf{EqGrph}$ is the graph with only one node, only one edge and with the single node equivalent only to itself, while the initial object is the graph with no nodes and no edges.
\end{example}

\begin{example}
    Given two graphs with equivalences $\mathcal{G} = (G, \sim_G) $ and $\mathcal{H} = (H, \sim_{H})$, one can construct the product graph $\mathcal{G \times H} = (G \times H, \sim_{G \times H})$, together with the two projections $\pi_G$ and $\pi_H$,  where $\sim_{G \times H}$ is the least equivalence relation on $G \times H$ such that, if $w_1$ and $w_2$ are nodes of $\mathcal{G \times H}$, we have $w_1 \sim_{G \times H} w_2$ if $\pi_G(w_1) \sim_G \pi_G(w_2)$ and $\pi_H(w_1) \sim_H \pi_H(w_2)$.
\end{example}

\begin{example}
    An equalizer in $\mathbf{EgGrph}$ for a pair of morphisms $f, g: \mathcal{G \rightarrow H}$ is exactly the same as in $\mathbf{Graph}$, namely, $(\mathcal{E} = (E, \sim_E), e)$, where $E$ is the greatest subgraph of the underlying graph of $\mathcal{G}$ on which $f$ and $g$ are equal, together with the inclusion $e: \mathcal{E \rightarrow G}$. The equivalence relation is preserved ($\sim_E$ is $\sim_G$ restricted to the set of vertices of $E$).
\end{example}

The two previous examples shows that the category $\mathbf{EgGrph}$ has both equalizer and finite products, so, by Theorem~\ref{th:limit}, we can conclude that $\mathbf{EgGrph}$ has finite limits.

\begin{example}
    Given two parallel morphisms $f, g: \mathcal{G \rightarrow H}$ in $\mathbf{EgGrph}$, the coequalizer $(\mathcal{C}, c)$ is the graph 
\end{example}
